---
title: "London Stock Pirce"
author: "Shi Fan Jin"
date: "11/11/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(astsa)
library(forecast)
```


##Your task is to predict the next 10 observations of these time series.
##You need to turn in the following:
###1. Your predictions for all three data sets (see below for further instructions).
###2. A report describing your analysis of one of the data sets (see below for further instructions).

```{r}
dat1 = read.csv("data1.csv")
dat2 = read.csv("data2.csv")
dat3 = read.csv("data3.csv")
```

#DATA 1
```{r}
plot(dat1, type = "l", main = "data 1")
```

The plot of the data set doesn't seem to be stationary, it looks like it has a seasonal factor and some kind of trend.
In this case, I decided difference the data with different lags in order to get rid of the seasonal factor. 

##Data 1 Transformation
```{r}
plot(sqrt(dat1$x), type = "l", main = "sqrt(dat1)")
# Nothing helpful

# Since there's around 3 peaks in every 100 period, try differencing with lag = 100/3 = 33.33, which lag = 33
dat1_lag33 = diff(dat1$x, lag = 33)
plot(dat1_lag33, type = "l", main = "Differenced data 1 with lag = 33")

auto.arima(dat1$x)
```


##Data Transforming (TRY)
```{r}
dat1_diff4 = diff(dat1$x, differences = 4)
plot(dat1_diff4, type = "l", main = "Differenced 4 times to remove quartic trend")
acf2(dat1_diff4)


# log transform                   # bad
dat1_log = log(dat1$x)
plot(dat1_log, type = "l", main = "log(dat1)")

#lag = 1                    # looks good
dat1_lag1 = diff(dat1$x, lag = 1)
plot(dat1_lag1, type = "l", main = "Differenced data 1 with lag = 1")

# Since the plot suggests a quartic trend
# Use 4th differencing to handle it
#lag = 4                    # looks good
dat1_lag4 = diff(dat1$x, lag = 4)
plot(dat1_lag4, type = "l", main = "Differenced data 1 with lag = 4")

# LOOKS NOT THAT GOOD
dat1_lag4 = diff(dat1$x, lag = 4)
plot(dat1_lag4, type = "l", main = "Differenced data 1 with lag = 4")
plot(diff(dat1_lag4, lag = 25), type = "l", main = "diff(diff(dat1, lag = 4), lag = 25)")

#lag = 10                   # looks good
dat1_lag10 = diff(dat1$x, lag = 10)
plot(dat1_lag10, type = "l", main = "Differenced data 1 with lag = 10")
# The first difference of the seasonal(10) differenced of data 
dat_lag10_diff = diff(dat1_lag10)
plot(dat_lag10_diff, type = "l", main = "diff(diff(dat1, lag = 10), lag = 1)")
# This looks like a white noise
dat_lag10_diff33 = diff(dat1_lag10, lag = 33)
plot(dat_lag10_diff33, type = "l", main = "diff(diff(dat1, lag = 10), lag = 33)")


#lag = 30                   # bad
dat1_lag30 = diff(dat1$x, lag = 30)
plot(dat1_lag30, type = "l", main = "Differenced data 1 with lag = 30")

#lag = 50                   # bad
dat1_lag50 = diff(dat1$x, lag = 50)
plot(dat1_lag50, type = "l", main = "Differenced data 1 with lag = 50")

# The first difference of the seasonal(50) differenced of data 
dat_lag50_diff = diff(dat1_lag50)
plot(dat_lag50_diff, type = "l", main = "diff(diff(dat1, lag = 50), lag = 1)")
# This looks like a white noise

# There seems to be 8 little peeks in one 100 period
plot(diff(diff(dat1$x), lag = 12),type = "l", main = "Differenced data 1 with lag = 12")
# still looks like some small seasonal components
# MAYBE DIFFERENCE AGAIN?
# The first difference of the seasonal(12) differenced of data
dat_lag12_diff = diff(diff(dat1$x), lag = 12)
plot(dat_lag12_diff, type = "l", main = "diff(diff(dat1$x), lag = 12)")
```

**CANDIDATES:**
dat1_diff4:
  diff(dat1$x, differences = 4)
  # 4 times differenced
dat1_lag1: 
  diff(dat1, lag = 1)
  # Differenced of lag(1)
dat_lag10_diff: 
  diff(diff(dat1, lag = 10), lag = 1)
  # The first difference of the seasonal(10) differenced of data 
dat_lag50_diff: 
  diff(diff(dat1, lag = 50), lag = 1)
  # The first difference of the seasonal(50) differenced of data 
dat_lag12_diff:
  diff(diff(dat1$x), lag = 12)
  # The first difference of the seasonal(12) differenced of data
  
```{r}
dat1_diff4 = diff(dat1$x, differences = 4)
plot(dat1_diff4, type = "l", main = "Differenced 4 times to remove quartic trend")

dat1_lag1 = diff(dat1$x, lag = 1)
plot(dat1_lag1, type = "l", main = "Differenced data 1 with lag = 1")

dat_lag10_diff = diff(dat1_lag10)
plot(dat_lag10_diff, type = "l", main = "diff(diff(dat1, lag = 10), lag = 1)")

dat_lag50_diff = diff(dat1_lag50)
plot(dat_lag50_diff, type = "l", main = "diff(diff(dat1, lag = 50), lag = 1)")

dat_lag12_diff = diff(diff(dat1$x), lag = 12)
plot(dat_lag12_diff, type = "l", main = "diff(diff(dat1$x), lag = 12)")
```

  
  
###Check the candidates if stationary using ACF
```{r}
acf2(dat1_diff4)
# potential p = 10, q = 4
# potential combination of (p, q) = (10, 0), (0, 4), (10, 4)
# HOWEVER, the PACF seems to be tapering, so in this case maybe it indicates a SMA(4)


acf2(dat1_lag1)
# NO SHARP CUT OFF IN THE BEGGINING >>> ignore the ones that's already in the beginning?
# potential p = 7, q = 7
# potential combination of (p, q) = (7, 0), (0, 7), (7, 7)
# HOWEVER, the PACF seems to be tapering, so in this case maybe it indicates a SMA(7)
# NO SECOND DIFF WITH SEASONAL LAG >>> CANT USE SAR OR SMA!!!


acf2(dat_lag10_diff)
acf(dat_lag10_diff)
# NO SHARP CUT OFF IN THE BEGGINING >>> ignore the ones that's already in the beginning?
# potential p = 2, q = 2
# potential combination of (p, q) = (2, 0), (0, 2), (2, 2)
# HOWEVER, the ACF seems to be tapering, so in this case maybe it indicates a SAR(2)


acf2(dat_lag50_diff)
acf(dat_lag50_diff)
pacf(dat_lag50_diff)
# NO SHARP CUT OFF IN THE BEGGINING >>> ignore the ones that's already in the beginning?
# potential p = 3, q = 4
# potential combination of (p, q) = (3, 0), (0, 4), (3, 4)
# HOWEVER, the ACF seems to be tapering, so in this case maybe it indicates a SAR(3)


acf2(dat_lag12_diff)
# NO SHARP CUT OFF IN THE BEGGINING >>> ignore the ones that's already in the beginning?
# potential p = 2, q = 2
# potential combination of (p, q) = (2, 0), (0, 2), (2, 2)
# HOWEVER, the ACF seems to be tapering, so in this case maybe it indicates a SAR(2)

```

##FITTING MODEL LATEST
```{r}
##################  FROM dat1_diff4
## potential combination of (p, q) = (10, 0), (0, 4), (10, 4)
#### (10,4,0)(0,0,0)[0]
model1A = sarima(dat1$x,p=10,d=4,q=0,P=0,D=0,Q=0,S=0)   #BAD
model1A$ttable
#### (0,4,4)(0,0,0)[0]
model1A_1 = sarima(dat1$x,p=0,d=4,q=4,P=0,D=0,Q=0,S=0)   #BAD
model1A_1$ttable
#### (10,4,4)(0,0,0)[0]
model1A_2 = sarima(dat1$x,p=10,d=4,q=4,P=0,D=0,Q=0,S=0)   #BAD
model1A_2$ttable


##################  FROM dat1_lag1
## potential combination of (p, q) = (7, 0), (0, 7), (7, 7)
#### (7,1,0)(0,0,0)[0]
model1B = sarima(dat1$x,p=7,d=1,q=0,P=0,D=0,Q=0,S=0)   #PRETTY GOOD
model1B$ttable
model1B$AIC
model1B$BIC
model1B$AICc
#### (0,1,7)(0,0,0)[0]
model1B_1 = sarima(dat1$x,p=0,d=1,q=7,P=0,D=0,Q=0,S=0)   #WORSE THAN PREVIOUS (model1B)
model1B_1$ttable
#### (7,1,7)(0,0,0)[0]
model1B_2 = sarima(dat1$x,p=7,d=1,q=7,P=0,D=0,Q=0,S=0)   #BAD
model1B_2$ttable
# FROM the p-values, MAYBE try AR(3) or/and MA(3)?
#### (3,1,3)(0,0,0)[0]
model1B_3 = sarima(dat1$x,p=3,d=1,q=3,P=0,D=0,Q=0,S=0)   #GOOD, AIC/BIC better than model1B, But BOX IS WORSE
model1B_3$ttable
model1B_3$AIC
model1B_3$BIC
model1B_3$AICc


##################  FROM dat_lag10_diff
## potential combination of (p, q) = (2, 0), (0, 2), (2, 2)
#### (2,1,0)(0,1,0)[10]
model1C = sarima(dat1$x,p=2,d=1,q=0,P=0,D=1,Q=0,S=10)   #BAD
model1C$ttable
#### (0,1,2)(0,1,0)[10]
model1C_1 = sarima(dat1$x,p=0,d=1,q=2,P=0,D=1,Q=0,S=10)   #BAD
model1C_1$ttable
#### (2,1,2)(0,1,0)[10]
model1C_2 = sarima(dat1$x,p=2,d=1,q=2,P=0,D=1,Q=0,S=10)   #BAD
model1C_2$ttable
model1C_2$AIC
model1C_2$BIC
model1C_2$AICc
# ACF SEEMS TO BE TAPERING >>> MAYBE SAR(2)?
#### (2,1,2)(2,1,0)[10]
model1C_3 = sarima(dat1$x,p=2,d=1,q=2,P=2,D=1,Q=0,S=10)   ##IMPROVING
model1C_3$ttable


##################  FROM dat_lag50_diff
## potential combination of (p, q) = (3, 0), (0, 4), (3, 4)
#### (3,1,0)(0,1,0)[50]
model1D = sarima(dat1$x,p=3,d=1,q=0,P=0,D=1,Q=0,S=50)   #BAD??? CANT REALLY TELL FROM THE BOX
model1D$ttable
model1D$AIC
model1D$BIC
model1D$AICc
# BAD AIC, BIC, AICc
#### (0,1,4)(0,1,0)[50]
model1D_1 = sarima(dat1$x,p=0,d=1,q=4,P=0,D=1,Q=0,S=50)   #THE AIC/BIC are BADDD, Although the BOX seems good in the FRONT
model1D_1$ttable
model1D_1$AIC
model1D_1$BIC
model1D_1$AICc
#### (3,1,4)(0,1,0)[50]
model1D_2 = sarima(dat1$x,p=3,d=1,q=4,P=0,D=1,Q=0,S=50)   #THE AIC/BIC are BADDD, Although the BOX seems good in the FRONT
model1D_2$ttable
model1D_2$AIC
model1D_2$BIC
model1D_2$AICc
# ACF SEEMS TAPERING >>> MAYBE SAR(3)?
#### (3,1,4)(3,1,0)[50]
#model1D_3 = sarima(dat1$x,p=3,d=1,q=4,P=3,D=1,Q=0,S=50)   #ERROR!



##################  FROM dat_lag12_diff
## potential combination of (p, q) = (2, 0), (0, 2), (2, 2)
#### (2,1,0)(0,1,0)[12]
model1E = sarima(dat1$x,p=2,d=1,q=0,P=0,D=1,Q=0,S=12)   #IMPROVING STILL BAD
model1E$ttable
#### (0,1,2)(0,1,0)[12]
model1E_1 = sarima(dat1$x,p=0,d=1,q=2,P=0,D=1,Q=0,S=12)   #IMPROVING STILL BAD
model1E_1$ttable
#### (2,1,2)(0,1,0)[12]
model1E_2 = sarima(dat1$x,p=2,d=1,q=2,P=0,D=1,Q=0,S=12)   #BAD
model1E_2$ttable
# ACF SEEMS TO BE TAPERING >>> MAYBE SAR(2)?
#### (2,1,2)(2,1,0)[12]
model1E_3 = sarima(dat1$x,p=2,d=1,q=2,P=2,D=1,Q=0,S=12)   #BAD
model1E_3$ttable
# OR MAYBE IT'S PACF TAPERING? >>> MAYBE SMA(2)?
#### (2,1,2)(0,1,2)[12]
model1E_4 = sarima(dat1$x,p=2,d=1,q=2,P=0,D=1,Q=2,S=12)   #PRETTY GOOD! BUT STILL SOME UNDER ## BIC BETTER THAN model1B_3 
model1E_4$ttable
model1E_4$AIC
model1E_4$BIC
model1E_4$AICc
## BIC BETTER THAN model1B_3, BUT AIC AND AICc a bit WORSE THAN model1B_3
# HOW ABOUT SAR(2) and SMA(2) at the same time?
#### (2,1,2)(2,1,2)[12]
model1E_5 = sarima(dat1$x,p=2,d=1,q=2,P=2,D=1,Q=2,S=12)   #BAD 
model1E_5$ttable


####################################################CHOOSE (3,1,3)(0,0,0)[0] OR (2,1,2)(0,1,2)[12]
model1B_3 = sarima(dat1$x,p=3,d=1,q=3,P=0,D=0,Q=0,S=0)
model1B_3$fit
model1B_3$AIC
model1B_3$BIC
model1B_3$AICc
# AIC, AICc better

model1E_4 = sarima(dat1$x,p=2,d=1,q=2,P=0,D=1,Q=2,S=12)
model1E_4$fit
model1E_4$AIC
model1E_4$BIC
model1E_4$AICc
# BIC better

####### >>>> NEED more COMPARISOM >>> MAYBE Cross Validation?


```

##Forecasting
```{r}
#######################(3,1,3)(0,0,0)[0] model
pred1A = sarima.for(dat1$x, n.ahead = 10, p=3,d=1,q=3,P=0,D=0,Q=0,S=0)
pred1A$pred

#######################(2,1,2)(0,1,2)[12] model ###############################USE THIS ONE IN THE END
pred1B = sarima.for(dat1$x, n.ahead = 10, p=2,d=1,q=2,P=0,D=1,Q=2,S=12)
pred1B$pred

```




###########################################################################################################################


#Data 2
```{r}
plot(dat2, type = "l", main = "data 2")
#Seems to have a seasonal factor and some kind of trend (quadratic, cubic)
```

##Data 2 Transformation
```{r}

# looks like there is a quadratic trend, take diff with differences = 2
dat2_diff2 = diff(dat2$x, differences = 2)
plot(dat2_diff2, type = "l", main = "Diff(dat2, differences = 2) to remove the quadratic trend")

plot(diff(diff(dat2$x)), type = "l", main = "diff(diff(dat2))")

##**diff(diff(dat2$x)) == diff(dat2$x, differences = 2)

# differenced with lag = 9
dat2_lag9 = diff(dat2$x, lag = 9)
plot(dat2_lag9, type = "l")
# there seems to be still a seasonal factor
dat2_lag9_10 = diff(dat2_lag9, lag = 10)
plot(dat2_lag9_10, type = "l")
# after differencing again with lag = 10


# looks like there's 9 peeks between 0 ~ 100, so each cycle would be 100/9,
# which is approximately 11.11, so choose lag = 11 for differencing for a try.
dat2_lag11 = diff(dat2$x, lag = 11)
plot(dat2_lag11, type = "l", main = "Differenced data 2 with lag = 11")
# This looks roughly stationary, with no seasonal component or linear trend. 
# To recap, dat2_lag11 = (1 - B^11)x_t where x_t is our original data. 

# looks like there's 8 peeks in each 100 period, cycle would be 100/8 = 12.5
# Try lag = 12
dat2_lag12 = diff(dat2$x, lag = 12)
plot(dat2_lag12, type = "l", main = "Differenced data 2 with lag = 12")
# This looks roughly stationary, with no seasonal component or linear trend. 
# To recap, dat2_lag11 = (1 - B^12)x_t where x_t is our original data. 
dat2_lag12_diff = diff(dat2_lag12)
plot(dat2_lag12_diff, type = "l", main = "diff(diff(dat2$x, lag = 12), lag = 1)")
# This looks roughly stationary, with no seasonal component or linear trend. 
# To recap, dat2_lag11 = (1 - B^12)(1 - B)x_t where x_t is our original data. 
```

**CANDIDATES:**
dat2_lag11: 
  diff(dat2, lag = 11)
  # Differenced of lag(1)
dat2_lag12: 
  diff(dat2, lag = 12)
  # Differenced of lag(12)
dat2_lag12_diff: 
  diff(diff(dat2, lag = 12), lag = 1)
  # The first difference of the seasonal(12) differenced of data 
```{r}
dat2_lag11 = diff(dat2$x, lag = 11)
plot(dat2_lag11, type = "l", main = "Differenced data 2 with lag = 11")

dat2_lag12 = diff(dat2$x, lag = 12)
plot(dat2_lag12, type = "l", main = "Differenced data 2 with lag = 12")

dat2_lag12_diff = diff(dat2_lag12)
plot(dat2_lag12_diff, type = "l", main = "diff(diff(dat2$x, lag = 12), lag = 1)")
```



###Check the candidates if stationary using ACF
```{r}
acf2(dat2_lag11) 
# potential p = 4, q = 2
# potential combination of (p, q) = (4, 0), (0, 2), (4, 2)
# HOWEVER, the PACF seems to be tapering, so in this case maybe it indicates a SMA(2)
# NO SECOND DIFF WITH SEASONAL LAG >>> CANT USE SAR OR SMA!!!

acf2(dat2_lag12)
# potential p = 1, q = 2
# potential combination of (p, q) = (1, 0), (0, 2), (1, 2)
# HOWEVER, the ACF seems to be tapering, so in this case maybe it indicates a SAR(1)
# NO SECOND DIFF WITH SEASONAL LAG >>> CANT USE SAR OR SMA!!!

acf2(dat2_lag12_diff)
# potential p = 8, q = 2
# potential combination of (p, q) = (8, 0), (0, 2), (8, 2)
# THIS ONE'S PACF DOESN'T SEEM TO BE GOOD (DOESN'T HAVE OBVIOUS CUTOFF IN THE FRONT)
# PACF SEEMS LIKE TAPPERING, MAYBE INCLUDE SMA(2)?
```

##Finding Fitted Model
```{r}
##################  FROM dat2_lag11
## potential combination of (p, q) = (4, 0), (0, 2), (4, 2)
#### (4,1,0)(0,0,0)[0]
model2A = sarima(dat2$x,p=4,d=1,q=0,P=0,D=0,Q=0,S=0)   #BAD
model2A$ttable
#### (0,1,2)(0,0,0)[0]
model2A_1 = sarima(dat2$x,p=0,d=1,q=2,P=0,D=0,Q=0,S=0)   #BAD
model2A_1$ttable
#### (4,1,2)(0,0,0)[0]
model2A_2 = sarima(dat2$x,p=4,d=1,q=2,P=0,D=0,Q=0,S=0)   #BAD
model2A_2$ttable



################## FROM dat2_lag12
## potential combination of (p, q) = (1, 0), (0, 2), (1, 2)
#### (1,1,0)(0,0,0)[0]
model2B = sarima(dat2$x,p=1,d=1,q=0,P=0,D=0,Q=0,S=0)   #BAD
model2B$ttable
#### (0,1,2)(0,0,0)[0]
model2B_1 = sarima(dat2$x,p=0,d=1,q=2,P=0,D=0,Q=0,S=0)   #IMPROVING
model2B_1$ttable
#### (1,1,2)(0,0,0)[0]
model2B_2 = sarima(dat2$x,p=1,d=1,q=2,P=0,D=0,Q=0,S=0)   #IMPROVING
model2B_2$ttable



##################  FROM dat2_lag12_diff
## potential combination of (p, q) = (8, 0), (0, 2), (8, 2)
#### (8,1,0)(0,1,0)[12]
model2C = sarima(dat2$x,p=8,d=1,q=0,P=0,D=1,Q=0,S=12)   #BAD
model2C$ttable
#### (0,1,2)(0,1,0)[12]
model2C_1 = sarima(dat2$x,p=0,d=1,q=2,P=0,D=1,Q=0,S=12)   #IMPROVING
model2C_1$ttable
#### (8,1,2)(0,1,0)[12]
model2C_2 = sarima(dat2$x,p=8,d=1,q=2,P=0,D=1,Q=0,S=12)   #BAD
model2C_2$ttable
# FROM THE p-values, suggested AR(1) or AR(2)? instead of AR(8)
#### (1,1,2)(0,1,0)[12]
model2C_3 = sarima(dat2$x,p=1,d=1,q=2,P=0,D=1,Q=0,S=12)   #IMPROVING
model2C_3$ttable
# FROM THE p-values, MAYBE MA(1) instead of MA(2)?
#### (1,1,1)(0,1,0)[12]
model2C_4 = sarima(dat2$x,p=1,d=1,q=1,P=0,D=1,Q=0,S=12)   #IMPROVING
model2C_4$ttable
model2C_4
# STILL NOT GOOD ENOUGH, MAYBE GO BACK TO ACF2 TO SEE IF THERE IS SEASONAL (P,D,Q)?
# PACF SEEMS LIKE TAPPERING >>> MAYBE INCLUDE SMA(2)?
#### (1,1,1)(0,1,2)[12]
model2C_5 = sarima(dat2$x,p=1,d=1,q=1,P=0,D=1,Q=2,S=12)   #VERY GOOD!
model2C_5$ttable
model2C_5$AIC
model2C_5$BIC
model2C_5$AICc

#### (1,1,1)(0,1,1)[12]
model2C_6 = sarima(dat2$x,p=1,d=1,q=1,P=0,D=1,Q=1,S=12)   #VERY GOOD!!!!    ######### BEST FOR NOW!!!!!
model2C_6$ttable
model2C_6$AIC
model2C_6$BIC
model2C_6$AICc

#### (1,1,1)(1,1,2)[12]
model2B_5 = sarima(dat2$x,p=1,d=1,q=1,P=1,D=1,Q=2,S=12)   #VERY GOOD!!!
model2B_5$ttable
model2B_5$AIC
model2B_5$BIC
model2B_5$AICc

################## FROM XEINAI
#### (1,1,1)(1,1,1)[12]
model2XN = sarima(dat2$x,p=1,d=1,q=1,P=1,D=1,Q=1,S=12)
model2XN$ttable
model2XN$AIC
model2XN$BIC
model2XN$AICc

model2TT = sarima(dat2$x,p=2,d=1,q=2,P=0,D=1,Q=2,S=12)
model2TT$ttable
# p-val suggest SMA(1)
model2TT_1 = sarima(dat2$x,p=1,d=1,q=2,P=0,D=1,Q=1,S=12)
model2TT_1$ttable
####################################################CHOOSE (1,1,1)(0,1,1)[12]
model2C_5$fit
```

##MODEL SELECTION
```{r}
#### (1,1,1)(0,1,1)[12]
#model2C_6 = sarima(dat2$x,p=1,d=1,q=1,P=0,D=1,Q=1,S=12)   #VERY GOOD!!!!    ######### BEST FOR NOW!!!!!
#model2C_6$ttable
model2C_6$AIC
model2C_6$BIC
model2C_6$AICc

#### (1,1,1)(1,1,2)[12]
#model2B_5 = sarima(dat2$x,p=1,d=1,q=1,P=1,D=1,Q=2,S=12)   #VERY GOOD!!!
#model2B_5$ttable
model2B_5$AIC
model2B_5$BIC
model2B_5$AICc

################## FROM XEINAI
#### (1,1,1)(1,1,1)[12]
#model2XN = sarima(dat2$x,p=1,d=1,q=1,P=1,D=1,Q=1,S=12)
#model2XN$ttable
model2XN$AIC
model2XN$BIC
model2XN$AICc


#######################LOOKING AT THE AIC/BIC/AICc, MODEL: (1,1,1)(0,1,1)[12] is the best! (Having the smallest values)
model2C_5 = sarima(dat2$x,p=1,d=1,q=1,P=0,D=1,Q=1,S=12)
```


##FORECASTING
```{r}
model2C_5

# USING (1,1,1)(0,1,1)[12]
pred2 <- sarima.for(xdata = dat2$x, n.ahead = 10, p=1,d=1,q=1,P=0,D=1,Q=1,S=12)
pred2$pred
plot(dat2$x, type = "l")

```


###########################################################################################################################

#Data 3
```{r}
plot(dat3, type = "l", main = "Data 3")
# looks like there are several seasonal factors
```


##Data 3 Transformation
```{r}
plot(log(dat3$x), type = "l", main = "log(dat3)")
# doesn't seem to be doing anything good

# Looks like there is some big seasonal factor which have length 200
# Difference with lag = 200 (the bigger seasonal factor)
dat3_lag200 = diff(dat3$x, lag = 200)
plot(dat3_lag200, type = "l", main = "Differenced data 3 with lag = 200")
# The mean looks roughly around 0, however, still some seasonal components
dat3_lag200_diff = diff(diff(dat3$x, lag = 200), lag = 1)
plot(dat3_lag200_diff, type = "l", main = "diff(diff(dat3$x, lag = 200), lag = 1)")

# looks like there's 10 peeks in each 100 period, cycle would be 100/10 = 10
# Try lag = 10
dat3_lag10 = diff(dat3$x, lag = 10)
plot(dat3_lag10, type = "l", main = "Differenced data 3 with lag = 10")
# This looks pretty good!
dat3_lag10_diff = diff(dat3_lag10, lag = 1)
plot(dat3_lag10_diff, type = "l", main = "diff(diff(dat3$x, lag = 10), lag = 1)")
# This looks even better (like white noise)


dat3_lag12 = diff(dat3$x, lag = 12)
plot(dat3_lag12, type = "l", main = "Differenced data 3 with lag = 12")
# Still have seasonal components
dat3_lag12_diff = diff(diff(dat3$x, lag = 12), lag = 1)
plot(dat3_lag12_diff, type = "l", main = "diff(diff(dat3$x, lag = 12), lag = 1)")
# Looks like stationary!

dat3_diff = diff(dat3$x)
plot(dat3_diff, type = "l", main = "diff(dat3$x)")
# Also looks like stationary!
```

**CANDIDATES:**
dat3_lag200_diff: 
  diff(diff(dat3$x, lag = 200), lag = 1)
  # The first difference of the seasonal(200) differenced of data 
dat3_lag10_diff: 
  diff(diff(dat3$x, lag = 10), lag = 1)
  # The first difference of the seasonal(10) differenced of data 
dat3_lag12_diff:
  diff(diff(dat3$x, lag = 12), lag = 1)
  # The first difference of the seasonal(12) differenced of data 
dat3_diff:
  diff(dat3$x)
  # Differenced data with lag = 1

###Check the candidates if stationary using ACF
```{r}
acf2(dat3_lag200_diff)
# potential p = 3, q = 2
# potential combination of (p, q) = (3, 0), (0, 2), (3, 2)
# HOWEVER, the PACF seems to be tapering, so in this case maybe it indicates a SMA(2)


acf2(dat3_lag10_diff)
# potential p = 3, q = 2
# potential combination of (p, q) = (3, 0), (0, 2), (3, 2)
# HOWEVER, the ACF seems to be tapering, so in this case maybe it indicates a SAR(3)
# MAYBE PACF is also TAPPERING? NOT SURE...


acf2(dat3_lag12_diff)
# potential p = 4, q = 4
# potential combination of (p, q) = (4, 0), (0, 4), (4, 4)
# HOWEVER, the PACF seems to be tapering, so in this case maybe it indicates a SMA(4)
# MAYBE ACF IS ALSO TAPPERING, SO MAYBE INCLUDE SAR(4)?


acf2(dat3_diff)
# potential p = 3, q = 4
# potential combination of (p, q) = (3, 0), (0, 4), (3, 4)
# HOWEVER, the PACF seems to be tapering, so in this case maybe it indicates a SMA(4)
# But MAYbe ACF also is tappering, in this case, maybe include a bit of SAR(3)
acf(dat3_diff)
pacf(dat3_diff)
```

## Model Fitting
```{r}
##################  FROM dat3_lag200_diff
## potential combination of (p, q) = (3, 0), (0, 2), (3, 2)
#### (3,1,0)(0,1,0)[200]
model3A = sarima(dat3$x,p=3,d=1,q=0,P=0,D=1,Q=0,S=200)   #BAD DOESN'T WORK! THERE'S AN ERROR
model3A$ttable
#### (0,1,2)(0,1,0)[200]
model3A_1 = sarima(dat3$x,p=0,d=1,q=2,P=0,D=1,Q=0,S=200)   #BAD DOESN'T WORK! THERE'S AN ERROR
model3A_1$ttable

## MAYBE LAG = 200 IS TOO BIG?


##################  FROM dat3_lag10_diff
## potential combination of (p, q) = (3, 0), (0, 2), (3, 2)
#### (3,1,0)(0,1,0)[10]
model3B = sarima(dat3$x,p=3,d=1,q=0,P=0,D=1,Q=0,S=10)   #IMPROVING
model3B$ttable
#### (0,1,2)(0,1,0)[10]
model3B_1 = sarima(dat3$x,p=0,d=1,q=2,P=0,D=1,Q=0,S=10)   #BAD
model3B_1$ttable
#### (3,1,2)(0,1,0)[10]
model3B_2 = sarima(dat3$x,p=3,d=1,q=2,P=0,D=1,Q=0,S=10)   #BAD
model3B_2$ttable
# ar1's p-value looks pretty good, MAYBE CHANGE TO AR(1)?
#### (1,1,2)(0,1,0)[10]
model3B_3 = sarima(dat3$x,p=1,d=1,q=2,P=0,D=1,Q=0,S=10)   #BAD
model3B_3$ttable
# ACF SEEMS TO BE TAPERING >>> SAR(3)?
#### (1,1,2)(3,1,0)[10]
model3B_4 = sarima(dat3$x,p=1,d=1,q=2,P=3,D=1,Q=0,S=10)   #A BIT IMPROVING
model3B_4$ttable
# FROM THE p-values, MAYBE try MA(1)?
#### (1,1,1)(3,1,0)[10]
model3B_5 = sarima(dat3$x,p=1,d=1,q=1,P=3,D=1,Q=0,S=10)   #BAD
model3B_5$ttable
# PACF ALSO? SEEMS TO BE TAPERING >>> SMA(2)?
#### (1,1,1)(3,1,2)[10]
model3B_6 = sarima(dat3$x,p=1,d=1,q=1,P=3,D=1,Q=2,S=10)   #BAD
model3B_6$ttable
# FROM THE p-values, Suggest SAR(1) and SMA(2)?
#### (1,1,1)(1,1,2)[10]
model3B_7 = sarima(dat3$x,p=1,d=1,q=1,P=1,D=1,Q=2,S=10)   #BAD
model3B_7$ttable
# FROM THE p-values, Suggest SAR(2)?
#### (1,1,1)(2,1,2)[10]
model3B_8 = sarima(dat3$x,p=1,d=1,q=1,P=2,D=1,Q=2,S=10)   #BAD
model3B_8$ttable


##################  FROM dat3_lag12_diff
## potential combination of (p, q) = (4, 0), (0, 4), (4, 4)
#### (4,1,0)(0,1,0)[12]
model3C = sarima(dat3$x,p=4,d=1,q=0,P=0,D=1,Q=0,S=12)   #BAD
model3C$ttable
#### (0,1,4)(0,1,0)[12]
model3C_1 = sarima(dat3$x,p=0,d=1,q=4,P=0,D=1,Q=0,S=12)   #BAD
model3C_1$ttable
#### (4,1,4)(0,1,0)[12]
model3C_2 = sarima(dat3$x,p=4,d=1,q=4,P=0,D=1,Q=0,S=12)   #BAD
model3C_2$ttable
# PACF SEEMS TO BE TAPERING >>> SMA(4)?
#### (4,1,4)(0,1,4)[12]
model3C_3 = sarima(dat3$x,p=4,d=1,q=4,P=0,D=1,Q=4,S=12)   #A BIT BETTER IN THE END?
model3C_3$ttable
# FROM THE p-values, Suggest AR(2)? MA(2) or MA(3)? SMA(1) or SMA(2)?
#### (2,1,2)(0,1,1)[12]
model3C_4 = sarima(dat3$x,p=2,d=1,q=2,P=0,D=1,Q=1,S=12)   #A BIT IMPROVING? 
model3C_4$ttable
#### (2,1,3)(0,1,1)[12]
model3C_5 = sarima(dat3$x,p=2,d=1,q=3,P=0,D=1,Q=1,S=12)   #IMPROVING
model3C_5$ttable
# FROM THE p-values, Suggest NO AR?
#### (0,1,3)(0,1,1)[12]
model3C_6 = sarima(dat3$x,p=0,d=1,q=3,P=0,D=1,Q=1,S=12)   #WORSE
model3C_6$ttable
# FROM THE p-values, Suggest MA(1) or MA(2)?
#### (0,1,1)(0,1,1)[12]
model3C_7 = sarima(dat3$x,p=0,d=1,q=1,P=0,D=1,Q=1,S=12)   #WORSE
model3C_7$ttable
#### (0,1,2)(0,1,1)[12]
model3C_8 = sarima(dat3$x,p=0,d=1,q=2,P=0,D=1,Q=1,S=12)   #WORSE
model3C_8$ttable
# SEEMS LIKE THERE SHOULD BE AR INCLUDED, CANNOT DELETE IT!
# BACK TO model3C_5 results, and take a look at the acf2 again!
# BACK TO THE ACF2 PLOT, MAYBE ACF IS ALSO TAPPERING?
# MAYBE TRY SAR(4)?
#### (2,1,3)(4,1,1)[12]
model3C_9 = sarima(dat3$x,p=2,d=1,q=3,P=4,D=1,Q=1,S=12)   #IMPROVING, BUT STILL WORSE THEN model3C_5
model3C_9$ttable
# FROM THE p-values, MAYBE SAR(2)?
#### (2,1,3)(2,1,1)[12]
model3C_10 = sarima(dat3$x,p=2,d=1,q=3,P=2,D=1,Q=1,S=12)   #IMPROVING, BUT STILL WORSE THEN model3C_5
model3C_10$ttable
# FROM THE p-values, MAYBE AR(1) and MA(2)?
#### (1,1,2)(2,1,1)[12]
model3C_11 = sarima(dat3$x,p=1,d=1,q=2,P=2,D=1,Q=1,S=12)   #IMPROVING, BUT STILL WORSE THEN model3C_5
model3C_11$ttable
# FROM THE p-values, MAYBE MA(1)?
#### (1,1,1)(2,1,1)[12]
model3C_12 = sarima(dat3$x,p=1,d=1,q=1,P=2,D=1,Q=1,S=12)   #IMPROVING, BUT STILL WORSE THEN model3C_5
model3C_12$ttable
# SEEMS LIKE GOING NOWHERE! TRY auto.arima FOR SOME CLUE!

auto.arima(dat3$x)
# FROM auto.arima(dat3$x), MAYBE TRY ARIMA(1,0,2)?
#### (1,0,2)(0,0,0)[0]
model3C_13 = sarima(dat3$x,p=1,d=0,q=2,P=0,D=0,Q=0,S=0)   #BAD
model3C_13$ttable
# MAYBE ADD BACK THE P, D, Q, S part?
#### (1,0,2)(2,1,1)[12]
model3C_14 = sarima(dat3$x,p=1,d=0,q=2,P=2,D=1,Q=1,S=12)   #IMPROVING, BUT STILL WORSE THEN model3C_5
model3C_14$ttable
# NOT GOING ANYWHERE!

##################  AUTO.ARIMA
auto.arima(dat3$x)
# Suggested ARIMA(1,0,2)
auto.arima(dat3$x, stepwise = FALSE, approximation = FALSE)
# Suggested ARIMA(1,0,4)


##################  FROM dat3_diff
## potential combination of (p, q) = (3, 0), (0, 4), (3, 4)
#### (3,1,0)(0,0,0)[0]
model3D = sarima(dat3$x,p=3,d=1,q=0,P=0,D=0,Q=0,S=0)   #BAD 
model3D$ttable
#### (0,1,4)(0,0,0)[0]
model3D_1 = sarima(dat3$x,p=0,d=1,q=4,P=0,D=0,Q=0,S=0)   #PRETTY GOOD!!!! ALL ABOVE DASH ######### BEST FOR NOW!!!!!
model3D_1$ttable
model3D_1$AIC
model3D_1$BIC
model3D_1$AICc
# SEE IF WE CAN GET BETTER RESULT >>> KEEP TRYING
#### (3,1,4)(0,0,0)[0]
model3D_2 = sarima(dat3$x,p=3,d=1,q=4,P=0,D=0,Q=0,S=0)   #A BIT WORSE THAN model3D_1
model3D_2$ttable


################## FROM XEINAI                  #model3D_1 is still better
#### (1,1,3)(0,0,0)[0]
#model3XN = sarima(dat3$x,p=1,d=1,q=3,P=0,D=0,Q=0,S=0)
#model3XN$ttable
model3XN$AIC
model3XN$BIC
model3XN$AICc
#model3XN_1 = sarima(dat3$x,p=1,d=1,q=3,P=2,D=1,Q=1,S=12)
#model3XN_1$ttable
model3XN_1$AIC
model3XN_1$BIC
model3XN_1$AICc



####################################################CHOOSE (0,1,4)(0,0,0)[0]
#model3D_1 = sarima(dat3$x,p=0,d=1,q=4,P=0,D=0,Q=0,S=0) 
#model3D_1$fit
model3D_1$AIC
model3D_1$BIC
model3D_1$AICc
```

##FORECASTING
```{r}
# USING (0,1,4)(0,0,0)[0]
pred3 <- sarima.for(xdata = dat3$x, n.ahead = 10, p=0,d=1,q=4,P=0,D=0,Q=0,S=0)
pred3$pred
plot(dat3$x, type = "l")
```


